Individual and Differential Harm in Redistricting

### 2022-12-06 paper/harm.Rmd
AL appendx + pres figures

As discussed in Section \ref{subsec:racial}, plaintiffs had to satisfy the first *Gingles* criterion by submitting one or more remedial maps demonstrating that a new Black opportunity-to-elect district could be drawn while meeting all other redistricting requirements.
In *Merrill,* the plaintiff's expert, Moon Duchin, submitted four such remedial plans, one of which is shown in Figure \ref{fig:al-maps}(e).
These plans form a natural counterfactual $\tilde Q$ to use in analyzing the challenged districting plan (Figure \ref{fig:al-maps}(d)), especially since the federal district court ruled in favor of the plaintiffs and found that their remedial plans satisfied the first *Gingles* criterion.

@@

>>>>

As discussed in Section \ref{subsec:racial}, plaintiffs had to satisfy the first *Gingles* criterion by submitting one or more remedial maps demonstrating that a new Black opportunity-to-elect district could be drawn while meeting all other redistricting requirements.
In *Merrill,* the plaintiff's expert, Moon Duchin, submitted four such remedial plans, one of which is shown in Figure \ref{fig:al-maps}(e).
These plans form a natural counterfactual $\tilde Q$ to use in analyzing the challenged districting plan (Figure \ref{fig:al-maps}(d)), especially since the federal district court ruled in favor of the plaintiffs and found that their remedial plans satisfied the first *Gingles* criterion.
In Appendix \ref{app:alsampling}, we repeat the analysis performed here on a different counterfactual consisting of plans sampled with a redistricting simulation algorithm; the results are broadly similar.

@@

# Additional Sampling Analysis for Alabama {#app:alsampling}

Here, we perform the same analysis as in Section \ref{sec:alabama}, but using a randomly sampled set of plans as a counterfactual, rather than the four plans used in litigation.
Specifically, we use 2,000 plans provided by the 50-State Simulation Project [@mccartan2022fiftystates].
These plans are designed to discourage packing of Black voters and maintain at least the same number of majority-opportunity districts as the enacted 2020 plan.

```{r alabama-harm-sample, fig.width=6.5}
#| fig.cap: |
#|      On thel eft, the expected fraction of voters in each precinct which are
#|      harmed by the 2020 enacted congressional districts versus the sampled 
#|      counterfactual set.
#|      On the right, the expected number of harmed voters in each race/partisan 
#|      group, relative to the total number of such voters.
include_graphics("figures/al_harm_sample.pdf")
```

Figure \@ref(fig:alabama-harm-sample) summarizes the results, which are mostly similar to the results in in Section \ref{sec:alabama}.
Across the state, the estimated average harm is 6.9%, which translates to around 144,000 voters.
Of these harmed voters, around 56,000 are Black, while 88,000 are White.
Similarly, 73,000 are Democratic, while only 71,000 are Republican.
Normalizing by group sizes, the expected harm for a Black Democrat is 9.6%, for a Black Republican is 16.8%, for a White Democrat is 9.7%, and for a White Republican is 5.2%.
So the challenged plan differentially harms Democratic voters, but does not differentially harm Black Democrats versus White Democrats.
However, it it especially harms Black Republicans, who have an estimated differential harm of 7.2% versus Black Democrats and of 11.6% versus White Republicans.
